---
title: "F-ratio histogram for scRNA data with varying threshold for HVG detection"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/home/tema/work/skolkovo/fish_project/")
knitr::opts_chunk$set(echo=FALSE, message = FALSE)
```

```{r, results = 'hide'}
library(tidyverse)
library(Seurat)
set.seed(42)
```


```{r}
# TODO:
# - 
# - 
# - 
# - 
# - 
```

```{r}

```

```{r subsetting_blood_for_each_sample_and_HVGs}
load(file = "data/rna.integrated.filtered.after.clustering.obj.RData")
rna.integrated.cluster.blood <- subset(rna.integrated, idents = 0)
blood_cells_ids <- rna.integrated.cluster.blood@assays[["RNA"]]@counts@Dimnames[[2]]
# blood_cells_ids <- paste(blood_cells_ids, rep("-1", length(blood_cells_ids)), sep = "") 


stickleback.rna.comb <- readRDS(file = "data/stickleback.rna.seurat.samples.PURE_ENSEMBLE.rds")
rna.list <- SplitObject(stickleback.rna.comb, split.by = "orig.ident")
vst.list <- data.frame()
vst_one <- data.frame()
for (i in 1:length(rna.list)) {
    # cells_to_keep <- gsub("-1", "", rna.list[[i]]@assays[["RNA"]]@counts@Dimnames[[2]])
    cells_to_keep <- rna.list[[i]]@assays[["RNA"]]@counts@Dimnames[[2]]
    cells_to_keep <- subset(cells_to_keep, cells_to_keep %in% blood_cells_ids)
    rna.list[[i]] <- subset(rna.list[[i]], cells = cells_to_keep)
    rna.list[[i]] <- NormalizeData(rna.list[[i]], verbose = FALSE)
    rna.list[[i]] <- FindVariableFeatures(rna.list[[i]], selection.method = "vst", 
        nfeatures = 5000, verbose = FALSE)
    tmp.frame <- data.frame(sample_vst = rna.list[[i]]@assays[["RNA"]]@meta.features[["vst.variance.standardized"]],
                            sample_ids = rna.list[[i]]@assays[["RNA"]]@counts@Dimnames[[1]])
    tmp.frame <- subset(tmp.frame, sample_ids %in% rna.list[[i]]@assays[["RNA"]]@var.features)
    tmp.frame <- tmp.frame[order(-tmp.frame$sample_vst),]
    tmp.frame <- mutate(tmp.frame, ID = row_number())
    tmp.frame$sample <- paste0("sample_", i)
    if(length(vst_one) == 0){
      
      vst_one <- tmp.frame
    }
    else{
      vst_one <- rbind(vst_one, tmp.frame)
    }
    tmp.frame$sample <- NULL
    
    colrenames <- c(paste0("sample_", i, "_variance_std"), paste0("sample_", i, "_ids"), "ID")
    colnames(tmp.frame) <- colrenames
    if(length(vst.list) == 0){
      vst.list <- tmp.frame
    }
    else{
      vst.list <- merge(vst.list, tmp.frame, by = "ID")
    }
}
# write.table(vst.list, file = "data/top_2000_HVGs_for_each_sample.tsv", sep = "\t")
# head(vst.list, 50)
```

```{r}
hvg_union <- unique(c(vst.list$sample_1_ids,
               vst.list$sample_2_ids,
               vst.list$sample_3_ids,
               vst.list$sample_4_ids))

hvg_counts_sample_1 <- tibble::rownames_to_column(as.data.frame(rna.list[["stickleback.sample.1"]]@assays[["RNA"]]@counts[rownames(rna.list[["stickleback.sample.1"]]@assays[["RNA"]]@data) %in% hvg_union,]))
hvg_counts_sample_2 <- tibble::rownames_to_column(as.data.frame(rna.list[["stickleback.sample.2"]]@assays[["RNA"]]@counts[rownames(rna.list[["stickleback.sample.2"]]@assays[["RNA"]]@data) %in% hvg_union,]))
hvg_counts_sample_3 <- tibble::rownames_to_column(as.data.frame(rna.list[["stickleback.sample.3"]]@assays[["RNA"]]@counts[rownames(rna.list[["stickleback.sample.3"]]@assays[["RNA"]]@data) %in% hvg_union,]))
hvg_counts_sample_4 <- tibble::rownames_to_column(as.data.frame(rna.list[["stickleback.sample.4"]]@assays[["RNA"]]@counts[rownames(rna.list[["stickleback.sample.4"]]@assays[["RNA"]]@data) %in% hvg_union,]))

hvg_counts_saline <- merge(hvg_counts_sample_1, hvg_counts_sample_2, by="rowname")
hvg_counts_freshwater <- merge(hvg_counts_sample_3, hvg_counts_sample_4, by="rowname")
```

```{r}
hvg_counts_all <- merge(hvg_counts_saline, hvg_counts_freshwater, by="rowname")
# saveRDS(hvg_counts_all, file = "testing/hvg_counts_scRNA_for_F_test.rds")

saline_right_boundary <- length(hvg_counts_saline) - 1
freshwater_left_boundary <- saline_right_boundary + 1 - 1
freshwater_right_boundary <- saline_right_boundary + length(hvg_counts_freshwater) - 1 - 1

output <- apply(hvg_counts_all[,2:freshwater_right_boundary], 1, function(row){
  x <- row[1:saline_right_boundary]
  x <- x[x != 0 & is.finite(x)]
  y <- row[freshwater_left_boundary:(freshwater_right_boundary)]
  y <- y[y != 0 & is.finite(y)]
  if(length(x) > 10 & length(y) > 10){
    saline_var <- var(x)
    freshwater_var <- var(y)
    f.test.res <- var.test(x, y)
    # if(is.finite(f.test.res$estimate) & f.test.res$estimate == 0){
    #   f.test.res$estimate = 1 # ========================================================== ARTIFICIALLY CHANGING ZEROES TO 1
    # }
    return(data.frame(f_ratio = f.test.res$estimate, p_val = f.test.res$p.value, saline_var = saline_var, freshwater_var = freshwater_var))
  }else{
    return(data.frame(f_ratio = NA, p_val = NA, saline_var = NA, freshwater_var = NA))
  }
})
f.test.frame <- do.call(rbind,output)
rownames(f.test.frame) <- hvg_counts_all$rowname

```

```{r}
# row <- as.numeric(as.vector(hvg_counts_all[1,2:freshwater_right_boundary]))
#   # message(paste0("row = ", row))
#   x <- row[1:saline_right_boundary]
#   # x <- x[is.finite(x)] # ================= removed filtration of zeros
#   y <- row[freshwater_left_boundary:freshwater_right_boundary]
#   # y <- y[is.finite(y)]
#   if(length(x) > 10 & length(y) > 10){
#     f.test.res <- var.test(x, y)
#     if(f.test.res$p.value == 0){
#       print(x[x != 0])
#       print("=============")
#       print(y[y != 0])
#       break
#     }
#   }
```



### F-test for dispersion for blood cluster cells separately

```{r f_test_for_blood_cluster}
f.test.filtered.blood <- subset(f.test.frame, p_val < 0.05 & f_ratio > 0 & f_ratio != Inf & p_val != 0)
# head(f.test.filtered, 10)
# write.table(f.test.filtered.blood, file = "data/f.test.filtered.blood.tsv", sep = '\t')
f.test.filtered.blood
```

### F-test histogram for blood cluster
```{r}
# f.test.filtered.blood <- read.table(file = "data/f.test.filtered.blood.tsv", sep = "\t", header = TRUE)
p_f_ratio_blood <- ggplot(f.test.filtered.blood, aes(x=f_ratio))+
  geom_histogram(binwidth = 0.25) +
  geom_vline(aes(xintercept=median(f_ratio)), color="blue", linetype="dashed", size=1) +
  coord_cartesian(xlim = c(0, 5)) +
  theme_classic(base_size = 20) +
  xlab("F-ratio") +
  ylab("Number of genes") +
  labs(title = "marine/freshwater variance in genes\n in scRNA data with 5000 HVG per sample")

# svg(filename = "plots/for_article/f_1f_F_ratio_histogram_for_blood_cluster.svg")
# print(p_f_ratio_blood)
# dev.off()
p_f_ratio_blood
```

### F-test histogram for blood cluster NO FILTERING

```{r}
# f.test.filtered.blood <- read.table(file = "data/f.test.filtered.blood.tsv", sep = "\t", header = TRUE)
 f.test.frame %>% 
  filter(is.finite(f_ratio), p_val < 0.05, p_val > 0) %>% 
  # transmute(f_ratio = log2(f_ratio)) %>% 
  ggplot(aes(x=f_ratio))+
  geom_histogram(binwidth = 0.25, fill = "#a2b1ba", color = "#899ba7") +
  geom_vline(aes(xintercept=median(f_ratio)), color="red", linetype="dashed", size=1) +
  coord_cartesian(xlim = c(0, 5)) +
  theme_classic(base_size = 20) +
  xlab("F-ratio") +
  ylab("Number of genes") -> p_f_ratio_blood
  # labs(title = "marine/freshwater variance in genes\n in scRNA data with 5000 HVG per sample") 

# svg(filename = "plots/for_article/f_1f_F_ratio_histogram_for_blood_cluster.svg")
# print(p_f_ratio_blood)
# dev.off()
p_f_ratio_blood
```

```{r}
# f.test.filtered.blood <- read.table(file = "data/f.test.filtered.blood.tsv", sep = "\t", header = TRUE)
 f.test.frame %>% 
  filter(is.finite(f_ratio), f_ratio > 0) %>% 
  # transmute(f_ratio = log2(f_ratio)) %>% 
  ggplot(aes(x=f_ratio))+
  geom_histogram(binwidth = 0.25, fill = "#a2b1ba", color = "#899ba7") +
  geom_vline(aes(xintercept=median(f_ratio)), color="red", linetype="dashed", size=1) +
  coord_cartesian(xlim = c(0, 5)) +
  theme_classic(base_size = 20) +
  xlab("F-ratio") +
  ylab("Number of genes") -> p_f_ratio_blood
  # labs(title = "marine/freshwater variance in genes\n in scRNA data with 5000 HVG per sample") 

# svg(filename = "plots/for_article/f_1f_F_ratio_histogram_for_blood_cluster.svg")
# print(p_f_ratio_blood)
# dev.off()
p_f_ratio_blood
```


```{r}
apply(hvg_counts_all[,-1], 1, sum) -> hvg_counts_all_rowsum
apply(hvg_counts_all[,-1], 1, mean) -> hvg_counts_all_rowmean
# apply(hvg_counts_all[,-1], 1, function(x){ifelse(x)}) -> hvg_counts_all_zeros_count
rowSums(hvg_counts_all[,-1] == 0) -> hvg_counts_all_zeros_count
hvg_counts_metrics <- tibble(gene = hvg_counts_all$rowname,
                             sum_expression = hvg_counts_all_rowsum,
                             zero_count = hvg_counts_all_zeros_count,
                             mean_expression = hvg_counts_all_rowmean)
```

```{r}
 hvg_counts_metrics %>% 
  # filter(is.finite(f_ratio), f_ratio > 0) %>%
  transmute(mean_expression = log2(mean_expression)) %>%
  ggplot(aes(x=mean_expression))+
  geom_histogram( fill = "#a2b1ba", color = "#899ba7") + 
  # xlab("F-ratio") +
  xlab("Log2 mean expression") +
  labs(title = "Mean expression in 1432 HVGs used for F-test in scRNA")
```

```{r}
 hvg_counts_metrics %>% 
  # filter(is.finite(f_ratio), f_ratio > 0) %>%
  # transmute(zero_count = log2(zero_count)) %>%
  ggplot(aes(x=zero_count))+
  geom_histogram( fill = "#a2b1ba", color = "#899ba7", aes(y =log10(..count..))) + 
  # xlab("F-ratio") +
  xlab("Log2 zero_count") +
  labs(title = "Zero count in 1432 HVGs used for F-test in scRNA")
```

```{r fig.width=11}
hvg_counts_metrics %>% 
  filter(log2(mean_expression) > -12, zero_count < 16500) -> expression_filtered_HVGs

 f.test.frame %>% 
  rownames_to_column(var = "gene") %>% 
  filter(is.finite(f_ratio), gene %in% expression_filtered_HVGs$gene) %>% 
  # transmute(f_ratio = log2(f_ratio)) %>% 
  ggplot(aes(x=f_ratio))+
  geom_histogram(binwidth = 0.25, fill = "#a2b1ba", color = "#899ba7") +
  geom_vline(aes(xintercept=median(f_ratio)), color="red", linetype="dashed", size=1) +
  # geom_vline(aes(xintercept=mean(f_ratio)), color="blue", linetype="dashed", size=1) +
  coord_cartesian(xlim = c(0, 6)) +
  theme_classic(base_size = 20) +
  xlab("F-ratio") +
  ylab("Number of genes") +
  labs(title = "marine/freshwater variance in genes
in scRNA data with 2000 HVG per sample
filtration: log2(mean_expression) > -12,
zero_count < 16500 (out of 19161 cells)") -> p_f_ratio_blood

# svg(filename = "plots/for_article/f_1f_F_ratio_histogram_for_blood_cluster.svg")
# print(p_f_ratio_blood)
# dev.off()
p_f_ratio_blood
```

```{r fig.width=11}
hvg_counts_metrics %>% 
  filter(log2(mean_expression) > -10) -> expression_filtered_HVGs

 f.test.frame %>% 
  rownames_to_column(var = "gene") %>% 
  filter(is.finite(f_ratio), gene %in% expression_filtered_HVGs$gene) %>% 
  # transmute(f_ratio = log2(f_ratio)) %>% 
  ggplot(aes(x=f_ratio))+
  geom_histogram(binwidth = 0.25, fill = "#a2b1ba", color = "#899ba7") +
  geom_vline(aes(xintercept=median(f_ratio)), color="red", linetype="dashed", size=1) +
  # geom_vline(aes(xintercept=mean(f_ratio)), color="blue", linetype="dashed", size=1) +
  coord_cartesian(xlim = c(0, 6)) +
  theme_classic(base_size = 20) +
  xlab("F-ratio") +
  ylab("Number of genes") +
  labs(title = "marine/freshwater variance in genes
in scRNA data with 2000 HVG per sample
filtration: log2(mean_expression) > -10") -> p_f_ratio_blood

# svg(filename = "plots/for_article/f_1f_F_ratio_histogram_for_blood_cluster.svg")
# print(p_f_ratio_blood)
# dev.off()
p_f_ratio_blood
```

# Testing box plotting

## Direct variance comparison

```{r}
# hvg_counts_all <- merge(hvg_counts_saline, hvg_counts_freshwater, by="rowname")
# # saveRDS(hvg_counts_all, file = "testing/hvg_counts_scRNA_for_F_test.rds")
# 
# saline_right_boundary <- length(hvg_counts_saline) - 1
# freshwater_left_boundary <- saline_right_boundary + 1 - 1
# freshwater_right_boundary <- saline_right_boundary + length(hvg_counts_freshwater) - 1 - 1
# 
# output <- apply(hvg_counts_all[,2:freshwater_right_boundary], 1, function(row){
#   x <- row[1:saline_right_boundary]
#   x <- x[x != 0 & is.finite(x)]
#   y <- row[freshwater_left_boundary:(freshwater_right_boundary)]
#   y <- y[y != 0 & is.finite(y)]
#   if(length(x) > 10 & length(y) > 10){
#     return(data.frame(saline_var = var(x), freshwater_var = var(y)))
#   }else{
#     return(data.frame(saline_var = NA, freshwater_var = NA))
#   }
# })
# var.test.frame <- do.call(rbind,output)
# rownames(var.test.frame) <- hvg_counts_all$rowname

```


```{r}
 f.test.frame %>% 
  pivot_longer(c("saline_var", "freshwater_var"), names_to = "salinity", values_to = "variance") %>%
  ggplot(aes(x = salinity, y = variance)) +
    geom_boxplot() +
    scale_y_continuous(limits = c(0,0.15)) +
    labs(title = "scRNA Direct variance comparison in saline and freshwater") -> p2
png(filename = "plots/scRNA_direct_variance_comparison.png") 
p2
dev.off()
p2
```

## Boxplots with division by F-ratio >< 1

```{r}
 f.test.frame %>% 
  filter(is.finite(f_ratio)) -> f_test_frame_fininte 
  f_test_frame_fininte$sign <- apply(f_test_frame_fininte, 1 , function(x){ifelse(x > 1, "saline", "freshwater")})["f_ratio",]
  
 f_test_frame_fininte %>% 
   transform(f_ratio = abs(log2(f_ratio))) %>% 
  ggplot(aes(x = sign, y = f_ratio)) +
    geom_boxplot() +
    scale_y_continuous(limits = c(0,3)) +
    labs(title = "Fold change in f-ratio comparison in 
         saline(f_ratio > 1) and freshwater(f_ratio < 1)")
```


```{r}
 f.test.frame %>% 
  filter(is.finite(f_ratio)) -> f_test_frame_fininte 
  f_test_frame_fininte$sign <- apply(f_test_frame_fininte, 1 , function(x){ifelse(x > 1, "saline", "freshwater")})["f_ratio",]
  
 f_test_frame_fininte %>% 
   transform(f_ratio = abs(log2(f_ratio))) %>% 
   filter(p_val < 0.05, p_val > 0) %>% 
  ggplot(aes(x = sign, y = f_ratio)) +
    geom_boxplot() +
    scale_y_continuous(limits = c(0,3)) +
    ylab("log2(f_ratio)") +
    labs(title = "Fold change in f-ratio comparison in 
         saline(f_ratio > 1) and freshwater(f_ratio < 1)
         Filtered: p_val < 0.05, p_val > 0")
```
