---
title: "scATAC data analysis ATAC2 QC and clustering with Signac"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/home/tema/work/skolkovo/fish_project/")
knitr::opts_chunk$set(echo=FALSE, message = FALSE)
```

```{r, results = 'hide'}
library(tidyverse)
library(Signac)
library(Seurat)
library(GenomeInfoDb)
library(patchwork)
set.seed(42)
```

```{r}
counts <- Read10X_h5("data/scATAC/atac2/filtered_peak_bc_matrix.h5")
```

```{r}
metadata <- read.csv(
  file = "data/scATAC/atac2/singlecell.csv",
  header = TRUE,
  row.names = 1
)

atac2_assay <- CreateChromatinAssay(
  counts = counts,
  sep = c(":", "-"),
  genome = "gasAcu1",
  fragments = 'data/scATAC/atac2/fragments.tsv.gz',
  min.cells = 1
)
```



```{r}
## Computing hash

atac2 <- CreateSeuratObject(
  counts = atac2_assay,
  assay = 'peaks',
  project = 'fish_scatac2',
  meta.data = metadata
)
```

# Banding patterns depending on mononucleosomal/nucleosome-free ratio (NS) 
# Fragment histogram for chr1

```{r fig.width=9}
atac2 <- NucleosomeSignal(object = atac2)
atac2$nucleosome_group <- ifelse(atac2$nucleosome_signal > 0.5, 'NS > 0.5', 'NS < 0.5')
FragmentHistogram(object = atac2, group.by = 'nucleosome_group', region = 'groupI-1-10000000')
```
Outlires with high NS should have different banding pattern, but here they exibit pretty similar pattern (do not know why). 


```{r}
load(file = "data/scATAC/ga_granges_annotation.RData")
```


```{r}
annotations <- ga_tx_merged
# change to UCSC style 
# seqlevelsStyle(annotations) <- 'UCSC'
genome(annotations) <- "gasAcu1"

# add the gene information to the object
Annotation(atac2) <- annotations
```


```{r}
atac2 <- TSSEnrichment(atac2, fast = FALSE)
```

```{r}
atac2$high.tss <- ifelse(atac2$TSS.enrichment > 1.3, 'High', 'Low')
TSSPlot(atac2, group.by = 'high.tss') + NoLegend()
```
Outliers with Low TSS enrichment should exibit almoust no TSS pattern, but these do (and I am not sure why).


```{r fig.width=14}
atac2$pct_reads_in_peaks <- atac2$peak_region_fragments / atac2$passed_filters * 100
# atac2$blacklist_ratio <- atac2$blacklist_region_fragments / atac2$peak_region_fragments

VlnPlot(
  object = atac2,
  features = c('pct_reads_in_peaks', 'peak_region_fragments',
               'TSS.enrichment', 'nucleosome_signal'),
  pt.size = 0.1,
  ncol = 4
)
```

```{r}
atac2_light_filtering <- subset(
  x = atac2,
  subset = peak_region_fragments > 200 &
    peak_region_fragments < 1000 &
    pct_reads_in_peaks > 30 &
    # blacklist_ratio < 0.025 &
    nucleosome_signal < 0.45 &
    TSS.enrichment > 1.3
)
atac2_light_filtering
```
### Normalization and linear dimensional reduction

```{r}
atac2 <- RunTFIDF(atac2)
atac2 <- FindTopFeatures(atac2, min.cutoff = 'q0')
atac2 <- RunSVD(object = atac2)
DepthCor(atac2)
```
The first LSI component often captures sequencing depth (technical variation) rather than biological variation. If this is the case, the component should be removed from downstream analysis.  


## Clustering with light filtering of cells (peak_region_fragments > 200)

```{r}
atac2_light_filtering <- RunUMAP(
  object = atac2_light_filtering,
  reduction = 'lsi',
  dims = 2:30
)
atac2_light_filtering <- FindNeighbors(
  object = atac2_light_filtering,
  reduction = 'lsi',
  dims = 2:30
)
atac2_light_filtering <- FindClusters(
  object = atac2_light_filtering,
  algorithm = 3,
  resolution = 1.2,
  verbose = FALSE
)

DimPlot(object = atac2_light_filtering, label = TRUE) + NoLegend()
```
I tryied heavier filtering below since we can see huge intermixing of clusters, which indicates bad quality of cells.



## Clustering with heavy filtering of cells (peak_region_fragments > 450)
 
```{r}
atac2_heavy_filtering <- subset(
  x = atac2_light_filtering,
  subset = peak_region_fragments > 450 &
    peak_region_fragments < 1000 &
    pct_reads_in_peaks > 30 &
    # blacklist_ratio < 0.025 &
    nucleosome_signal < 0.45 &
    TSS.enrichment > 1.3
)
atac2_heavy_filtering
```

```{r fig.width=14}
atac2_heavy_filtering$pct_reads_in_peaks <- atac2_heavy_filtering$peak_region_fragments / atac2_heavy_filtering$passed_filters * 100
# atac2$blacklist_ratio <- atac2$blacklist_region_fragments / atac2$peak_region_fragments

VlnPlot(
  object = atac2_heavy_filtering,
  features = c('pct_reads_in_peaks', 'peak_region_fragments',
               'TSS.enrichment', 'nucleosome_signal'),
  pt.size = 0.1,
  ncol = 4,
  group.by = "orig.ident"
)
```


```{r}
atac2_heavy_filtering <- RunUMAP(
  object = atac2_heavy_filtering,
  reduction = 'lsi',
  dims = 2:30
)
atac2_heavy_filtering <- FindNeighbors(
  object = atac2_heavy_filtering,
  reduction = 'lsi',
  dims = 2:30
)
atac2_heavy_filtering <- FindClusters(
  object = atac2_heavy_filtering,
  algorithm = 3,
  resolution = 1.2,
  verbose = FALSE
)

DimPlot(object = atac2_heavy_filtering, label = TRUE) + NoLegend()
```
Additional filtering has produced much nicer cluster separation.


```{r}
save(atac2_heavy_filtering, file = "data/scATAC/atac2_heavy_filtering_obj_qc_clusters.RData")
```
