---
title: "scATAC data analysis atac3 QC and clustering with Signac"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(echo=FALSE, message = FALSE)
```

```{r, results = 'hide'}
library(tidyverse)
library(Signac)
library(Seurat)
library(GenomeInfoDb)
library(patchwork)
set.seed(42)
```


```{r}
counts_atac3 <- Read10X_h5("data/scATAC/atac3/filtered_peak_bc_matrix.h5")
```

```{r}
metadata_atac3 <- read.csv(
  file = "data/scATAC/atac3/singlecell.csv",
  header = TRUE,
  row.names = 1
)

atac3_assay <- CreateChromatinAssay(
  counts = counts_atac3,
  sep = c(":", "-"),
  genome = "gasAcu1",
  fragments = 'data/scATAC/atac3/fragments.tsv.gz',
  min.cells = 1
)
```



```{r}
## Computing hash

atac3 <- CreateSeuratObject(
  counts = atac3_assay,
  assay = 'peaks',
  project = 'fish_scatac3',
  meta.data = metadata_atac3
)
```


# Banding patterns depending on mononucleosomal/nucleosome-free ratio (NS) 
# Fragment histogram for chr1


```{r fig.width=9}
atac3 <- NucleosomeSignal(object = atac3)
atac3$nucleosome_group <- ifelse(atac3$nucleosome_signal > 0.5, 'NS > 0.5', 'NS < 0.5')
FragmentHistogram(object = atac3, group.by = 'nucleosome_group', region = 'groupI-1-10000000')
```
Outlires with high NS should have different banding pattern, but here they exibit pretty similar pattern (do not know why). 



```{r}
load(file = "data/scATAC/ga_granges_annotation.RData")
```


```{r}
annotations <- ga_tx_merged
# change to UCSC style 
# seqlevelsStyle(annotations) <- 'UCSC'
genome(annotations) <- "gasAcu1"

# add the gene information to the object
Annotation(atac3) <- annotations
```


```{r}
atac3 <- TSSEnrichment(atac3, fast = FALSE)
```

```{r}
atac3$high.tss <- ifelse(atac3$TSS.enrichment > 1.8, 'High', 'Low')
TSSPlot(atac3, group.by = 'high.tss') + NoLegend()
```
Outliers with Low TSS enrichment should exibit almoust no TSS pattern, but these do (and I am not sure why).



```{r fig.width=14}
atac3$pct_reads_in_peaks <- atac3$peak_region_fragments / atac3$passed_filters * 100
# atac3$blacklist_ratio <- atac3$blacklist_region_fragments / atac3$peak_region_fragments

VlnPlot(
  object = atac3,
  features = c('pct_reads_in_peaks', 'peak_region_fragments',
               'TSS.enrichment', 'nucleosome_signal'),
  pt.size = 0.1,
  ncol = 4
)
```

```{r}
atac3 <- subset(
  x = atac3,
  subset = peak_region_fragments > 500 &
    peak_region_fragments < 2000 &
    pct_reads_in_peaks > 30 &
    # blacklist_ratio < 0.025 &
    nucleosome_signal < 0.7 &
    TSS.enrichment > 1.5
)
atac3
```


### Normalization and linear dimensional reduction

```{r}
atac3 <- RunTFIDF(atac3)
atac3 <- FindTopFeatures(atac3, min.cutoff = 'q0')
atac3 <- RunSVD(object = atac3)
DepthCor(atac3)
```
The first LSI component often captures sequencing depth (technical variation) rather than biological variation. If this is the case, the component should be removed from downstream analysis.  


```{r}
atac3 <- RunUMAP(
  object = atac3,
  reduction = 'lsi',
  dims = 2:30
)
atac3 <- FindNeighbors(
  object = atac3,
  reduction = 'lsi',
  dims = 2:30
)
atac3 <- FindClusters(
  object = atac3,
  algorithm = 3,
  resolution = 0.9,
  verbose = FALSE
)

DimPlot(object = atac3, label = TRUE) + NoLegend()
```
No clear separation of the clusters, probably due to low cell count and poor sequencing quality (not sure why). Probably best to drop this sample.


```{r}
save(atac3, file = "data/scATAC/atac3_obj_qc_clusters.RData")
```

