---
title: "Estimating variance difference between pseudo-bulk data from blood cluster between samples scATAC. (modeling blood cluster heteroscedasticity between samples)"
date: "`r Sys.Date()`"
author: "Artemiy Golden"
output:
  rmdformats::robobook:
    code_folding: hide
    self_contained: true
    thumbnails: false
    lightbox: true
---


```{r setup, include=FALSE}
library(knitr)
## Global options
options(max.print = "75")
knitr::opts_chunk$set(
  echo = TRUE, cache = TRUE, prompt = FALSE,
  tidy = FALSE, comment = NA,
  message = FALSE, warning = FALSE
)
opts_knit$set(width = 75)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r, results = 'hide'}
library(tidyverse)
library(Seurat)
library(ggpubr)
library(venn)
library(clusterProfiler)
library(biomaRt)
library(enrichplot)
library(Signac)
library(limma)
library(edgeR)
library(scuttle)
```





```{r}
load(file = "data/scATAC/combined_scATAC_samples_obj.RData")

```

```{r}
# Creating downsampled object
minimum_cells_per_sample <- min(table(combined@meta.data[["sample"]]))
ratio_of_downsampling_compared_to_1_sample = 1

cell_id_sample <- tibble(
  id = names(Idents(combined)),
  sample = combined@meta.data[["sample"]])

cell_id_sample_downsampled <- cell_id_sample %>%
  dplyr::filter(sample != "atac3") %>% 
  dplyr::group_by(sample) %>%
  slice_sample(n = minimum_cells_per_sample * ratio_of_downsampling_compared_to_1_sample) %>% 
  full_join(y = cell_id_sample[cell_id_sample$sample == "atac3",])

obj_downsampled <- subset(combined, cells = cell_id_sample_downsampled$id)
```


```{r}
pseudo_bulk_by_sample <- sumCountsAcrossCells(combined@assays$ATAC@counts, combined@meta.data[["sample"]])
```

```{r}
DGEList(as.matrix(pseudo_bulk_by_sample@assays@data@listData[["sum"]])) -> y
y <- calcNormFactors(y)
y$samples$group <- pseudo_bulk_by_sample@colData@rownames

cd <- c("saline", "saline", "freshwater", "freshwater")
model.matrix(~0+cd) -> design
```

VoomByGroup function is used to estimate the variance per pseudo-bulk sample. It was taken from the following pre-print: https://www.biorxiv.org/content/biorxiv/early/2022/09/14/2022.09.12.507511.full.pdf
Source code for VoomByGroup was found at: https://github.com/YOU-k/voomByGroup
BCV - biological coefficient of variation


## No downsampling
 

```{r}
source("external/voomByGroup/voomByGroup.R")
voomByGroup(y,design = design, group = cd, plot = "combine") -> y_vbg
```

## Downsampled to the sample with minimal number of cells
 
 

```{r}
pseudo_bulk_by_sample <- sumCountsAcrossCells(obj_downsampled@assays$ATAC@counts, obj_downsampled@meta.data[["sample"]])
DGEList(as.matrix(pseudo_bulk_by_sample@assays@data@listData[["sum"]])) -> y
y <- calcNormFactors(y)
y$samples$group <- pseudo_bulk_by_sample@colData@rownames

cd <- c("saline", "saline", "freshwater", "freshwater")
model.matrix(~0+cd) -> design

source("external/voomByGroup/voomByGroup.R")
voomByGroup(y,design = design, group = cd, plot = "combine") -> y_vbg
```

## Downsampled, pseudo-bulk with 3 technical replicates per each sample
 
After downsampling each of the samples was randomly divided into 3 technical replicates.

```{r}
# Create pseudo-bulk replicates
ids <- tibble(sample = obj_downsampled@meta.data[["sample"]],
              replicate = rep(sample(c("_a", "_b", "_c"), length(sample) / 4, replace = TRUE), 4)) %>% 
  mutate(id = paste0(sample, replicate)) %>% 
  pull(id)
pseudo_bulk_by_sample <- sumCountsAcrossCells(obj_downsampled@assays$ATAC@counts, ids)

DGEList(as.matrix(pseudo_bulk_by_sample@assays@data@listData[["sum"]])) -> y
y <- calcNormFactors(y)
y$samples$group <- pseudo_bulk_by_sample@colData@rownames

cd <- c(rep("saline", 6), rep("freshwater", 6))
model.matrix(~0+cd) -> design

source("external/voomByGroup/voomByGroup.R")
voomByGroup(y,design = design, group = cd, plot = "combine") -> y_vbg
```






