---
title: "Estimating variance difference between pseudo-bulk data from blood cluster between samples. (modeling blood cluster heteroscedasticity between samples)"
date: "`r Sys.Date()`"
author: "Artemiy Golden"
output:
  rmdformats::robobook:
    code_folding: hide
    self_contained: true
    thumbnails: false
    lightbox: true
---


```{r setup, include=FALSE}
library(knitr)
## Global options
options(max.print = "75")
knitr::opts_chunk$set(
  echo = TRUE, cache = TRUE, prompt = FALSE,
  tidy = FALSE, comment = NA,
  message = FALSE, warning = FALSE
)
opts_knit$set(width = 75)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r, results = 'hide'}
library(tidyverse)
library(Seurat)
library(ggpubr)
library(venn)
library(clusterProfiler)
library(biomaRt)
library(enrichplot)
```

```{r}
# generate pseudo-bulk replicates from the blood cluster
# install voom
# voomByGroup function, by specifying plot=“separate” to get individual mean-variance plots for each group (see https://www.biorxiv.org/content/biorxiv/early/2022/09/14/2022.09.12.507511.full.pdf)

```



```{r}
pbmc_pb <- readRDS(file = "external/voomByGroup/data/pbmc_pb.rds")
load(file = "data/rna.integrated.filtered.after.clustering.obj.RData")
rna.integrated.cluster.blood <- subset(rna.integrated, idents = 0)
```

```{r}
# Creating downsampled object
minimum_cells_per_sample <- min(table(rna.integrated.cluster.blood@meta.data[["orig.ident"]]))
ratio_of_downsampling_compared_to_1_sample = 1

cell_id_sample <- tibble(
  id = names(Idents(rna.integrated.cluster.blood)),
  sample = rna.integrated.cluster.blood@meta.data[["orig.ident"]])

cell_id_sample_downsampled <- cell_id_sample %>%
dplyr::filter(sample != "stickleback.sample.1") %>% 
dplyr::group_by(sample) %>%
slice_sample(n = minimum_cells_per_sample * ratio_of_downsampling_compared_to_1_sample) %>% 
full_join(y = cell_id_sample[cell_id_sample$sample == "stickleback.sample.1",])

blood_obj_downsampled <- subset(rna.integrated.cluster.blood, cells = cell_id_sample_downsampled$id)
```


```{r}
require(scuttle)
pseudo_bulk_by_sample <- sumCountsAcrossCells(rna.integrated.cluster.blood@assays$RNA@counts, rna.integrated.cluster.blood@meta.data[["orig.ident"]])
```

```{r}
library(limma)
library(edgeR)
```

```{r}
DGEList(as.matrix(pseudo_bulk_by_sample@assays@data@listData[["sum"]])) -> y
y <- calcNormFactors(y)
y$samples$group <- pseudo_bulk_by_sample@colData@rownames

cd <- c("saline", "saline", "freshwater", "freshwater")
model.matrix(~0+cd) -> design
```

VoomByGroup function is used to estimate the variance per pseudo-bulk sample. It was taken from the following pre-print: https://www.biorxiv.org/content/biorxiv/early/2022/09/14/2022.09.12.507511.full.pdf
Source code for VoomByGroup was found at: https://github.com/YOU-k/voomByGroup
BCV - biological coefficient of variation


## No downsampling
 

```{r}
source("external/voomByGroup/voomByGroup.R")
voomByGroup(y,design = design, group = cd, plot = "combine") -> y_vbg
```

## Downsampled to the sample with minimal number of cells
 
 

```{r}
pseudo_bulk_by_sample <- sumCountsAcrossCells(blood_obj_downsampled@assays$RNA@counts, blood_obj_downsampled@meta.data[["orig.ident"]])
DGEList(as.matrix(pseudo_bulk_by_sample@assays@data@listData[["sum"]])) -> y
y <- calcNormFactors(y)
y$samples$group <- pseudo_bulk_by_sample@colData@rownames

cd <- c("saline", "saline", "freshwater", "freshwater")
model.matrix(~0+cd) -> design

source("external/voomByGroup/voomByGroup.R")
voomByGroup(y,design = design, group = cd, plot = "combine") -> y_vbg
```

## Downsampled, pseudo-bulk with 3 technical replicates per each sample
 
After downsampling each of the samples was randomly divided into 3 technical replicates.

```{r}
# Create pseudo-bulk replicates
ids <- tibble(sample = blood_obj_downsampled@meta.data[["orig.ident"]],
              replicate = rep(sample(c("_a", "_b", "_c"), length(sample) / 4, replace = TRUE), 4)) %>% 
  mutate(id = paste0(sample, replicate)) %>% 
  pull(id)
pseudo_bulk_by_sample <- sumCountsAcrossCells(blood_obj_downsampled@assays$RNA@counts, ids)

DGEList(as.matrix(pseudo_bulk_by_sample@assays@data@listData[["sum"]])) -> y
y <- calcNormFactors(y)
y$samples$group <- pseudo_bulk_by_sample@colData@rownames

cd <- c(rep("saline", 6), rep("freshwater", 6))
model.matrix(~0+cd) -> design

source("external/voomByGroup/voomByGroup.R")
voomByGroup(y,design = design, group = cd, plot = "combine") -> y_vbg
```






