---
title: "Histograms of pairwise distance saline/freshwater and chi-square test for phenotype distribution in clusters"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/home/tema/work/skolkovo/fish_project/")
knitr::opts_chunk$set(echo=FALSE, message = FALSE)
```

```{r, results = 'hide'}
library(ggplot2)
library(dplyr)
library(Seurat)
```

```{r separating_blood_cluster}
load(file = "data/rna.integrated.filtered.after.clustering.obj.RData")
rna.integrated.cluster.blood <- subset(rna.integrated, idents = 0)
rna.integrated.cluster.rest <- subset(rna.integrated, idents = c(1,2))
```

### PCA plot of the blood cluster by the samples

```{r blood_cluster_PCA_plot}
rna.integrated.cluster.blood <- FindNeighbors(rna.integrated.cluster.blood, dims = 1:15)
rna.integrated.cluster.blood  <- FindClusters(rna.integrated.cluster.blood, resolution = 0.3)
rna.integrated.cluster.blood <- RunPCA(rna.integrated.cluster.blood)
DimPlot(rna.integrated.cluster.blood, reduction = "pca", group.by = "orig.ident")
```

```{r calculate_distances_from_pca}
set.seed(42)
pca_raw <- as.data.frame(rna.integrated.cluster.blood@reductions[["pca"]]@cell.embeddings)
pca_raw$sample <- rna.integrated.cluster.blood@meta.data[["orig.ident"]]
pca_raw <- pca_raw[sample(nrow(pca_raw)), c(1:10,51)]
pca_saline <- subset(pca_raw, sample %in% c("stickleback.sample.1", "stickleback.sample.2"))[,1:10]
pca_freshwater <- subset(pca_raw, sample %in% c("stickleback.sample.3", "stickleback.sample.4"))[,1:10]

pca_saline_1_half <- pca_saline[1:4672,]
pca_saline_2_half <- pca_saline[4674:nrow(pca_saline),]
colnames(pca_saline_2_half) <- paste(colnames(pca_saline_2_half), ".2half", sep="")
pca_salineVsaline <- cbind(pca_saline_1_half, pca_saline_2_half)
dist_matrix_salineVsaline <- fields::rdist(pca_saline_2_half, pca_saline_2_half)

dist_matrix_salineVfreshwater <- fields::rdist(pca_saline, pca_freshwater[1:nrow(pca_saline),])

pca_freshwater_1_half <- pca_freshwater[1:4907,]
pca_freshwater_2_half <- pca_freshwater[4909:nrow(pca_freshwater),]
colnames(pca_freshwater_2_half) <- paste(colnames(pca_freshwater_2_half), ".2half", sep="")
pca_freshwaterVfreshwater <- cbind(pca_freshwater_1_half, pca_freshwater_2_half)
dist_matrix_freshwaterVfreshwater <- fields::rdist(pca_freshwater_2_half, pca_freshwater_2_half)


# dist_salineVsaline <- apply(pca_salineVsaline, Eucle)
```

### Histograms of pairwise distance saline/freshwater

```{r parwise_histrograms_of_distances_between_diff_water_types}
frame_for_hist <- data.frame(dist = c(as.vector(dist_matrix_salineVsaline), as.vector(dist_matrix_salineVfreshwater), as.vector(dist_matrix_freshwaterVfreshwater)),
                             pair = c(rep("salineVsaline", length(dist_matrix_salineVsaline)),
                                      rep("salineVfreshwater", length(dist_matrix_salineVfreshwater)),
                                      rep("freshwaterVfreshwater", length(dist_matrix_freshwaterVfreshwater))))

mu <- plyr::ddply(frame_for_hist, "pair", summarise, grp.mean=mean(dist))

small_frame_for_hist <- frame_for_hist[sample(nrow(frame_for_hist), 1000000),]

p <- ggplot(small_frame_for_hist, aes(x=dist, fill=pair, color=pair)) +
  geom_histogram(position="identity", alpha=0.5, binwidth = 0.25) + 
  geom_vline(data=mu, aes(xintercept=grp.mean, color=pair),
             linetype="dashed")
p
```




```{r salineVsaline_histogram}
# hist(dist_matrix_salineVsaline, breaks=168)
```

```{r salineVfreshwater_histogram}
# hist(dist_matrix_salineVfreshwater, breaks=168)
```

```{r freshwaterVfreshwater_histogram}
# hist(dist_matrix_freshwaterVfreshwater, breaks=84)
```

### UMAP plot of clusters excluding blood

```{r non_blood_cells_umap}
#=========================== Rest cluster UMAP plot  =================================
rna.integrated.cluster.rest <- FindNeighbors(rna.integrated.cluster.rest, dims = 1:15)
rna.integrated.cluster.rest  <- FindClusters(rna.integrated.cluster.rest, resolution = 0.15)
rna.integrated.cluster.rest <- RunUMAP(rna.integrated.cluster.rest, dims = 1:15)
DimPlot(rna.integrated.cluster.rest, reduction = "umap")
```

### PCA plot of clusters excluding blood

```{r non_blood_cells_pca}
rna.integrated.cluster.rest <- RunPCA(rna.integrated.cluster.rest)
DimPlot(rna.integrated.cluster.rest, reduction = "pca")
```

### Chi-square test for the distribution of saline/freshwater across samples

```{r chi_square_test_for_sample_distribution_in_clusters}
samples_over_clusters <- data.frame(sample = rna.integrated.cluster.rest@meta.data[["orig.ident"]],
                                    cluster = as.character(rna.integrated.cluster.rest@active.ident))
water_sample_table <- data.frame(sample = unique(rna.integrated.cluster.rest@meta.data[["orig.ident"]]),
                                 water = c("saline", "saline", "freshwater", "freshwater"))
samples_over_clusters <- merge(samples_over_clusters, water_sample_table, by = "sample")

# normalizing for the difference in cell numbers salineVfreshwater for the whole run
correction_coefficient = nrow(pca_freshwater) / nrow(pca_saline)

table_as_frame <- as.data.frame(samples_over_clusters_table)
samples_over_clusters_matrix <- as.data.frame(matrix(table_as_frame$Freq, ncol=2, dimnames = list(unique(table_as_frame$Var1), unique(table_as_frame$Var2))))

samples_over_clusters_matrix$freshwater <- samples_over_clusters_matrix$freshwater / correction_coefficient
as.matrix(samples_over_clusters_matrix)

chisq.test(samples_over_clusters_matrix, correct = FALSE)
```

Frequencies of cells have been normalized to adjust for different number of sequenced freshwater and saline cells. The table above shows normalized numbers of cells.


