---
title: "Processing scRNA-seq dataset of stickleback fish"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/home/tema/work/skolkovo/fish_project/data/")
knitr::opts_chunk$set(echo=FALSE, message = FALSE)
```

```{r, results = 'hide'}
library(ggplot2)
library(dplyr)
library(Seurat)
```

```{r}
#=============================================== SETUP =============================================
# Merging all samples into one seurat object
cdna1.data <- Read10X(data.dir = "cdna1_counts/", gene.column = 2)
cdna2.data <- Read10X(data.dir = "cdna2_counts/", gene.column = 2)
cdna3.data <- Read10X(data.dir = "cdna3_counts/", gene.column = 2)
cdna4.data <- Read10X(data.dir = "cdna4_counts/", gene.column = 2)


```

## Sample 1 QC

```{r fig.width=11}
cdna1 <- CreateSeuratObject(counts = cdna1.data, project = "stickleback.sample.1")
cdna1[["percent.mt"]] <- PercentageFeatureSet(cdna1, pattern = "^mt-"  )
cdna1 <- subset(cdna1, subset = nFeature_RNA > 100 & nFeature_RNA < 500 & percent.mt < 2)
VlnPlot(cdna1, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, pt.size = 0.1) 
plot1 <- FeatureScatter(cdna1, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(cdna1, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
CombinePlots(plots = list(plot1, plot2))
```

## Sample 2 QC

```{r  fig.width=11}
cdna2 <- CreateSeuratObject(counts = cdna2.data, project = "stickleback.sample.2")
cdna2[["percent.mt"]] <- PercentageFeatureSet(cdna2, pattern = "^mt-"  )
cdna2 <- subset(cdna2, subset = nFeature_RNA > 100 & nFeature_RNA < 500 & percent.mt < 2)
VlnPlot(cdna2, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, pt.size = 0.1) 
plot1 <- FeatureScatter(cdna2, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(cdna2, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
CombinePlots(plots = list(plot1, plot2))
```

## Sample 3 QC

```{r  fig.width=11}
cdna3 <- CreateSeuratObject(counts = cdna3.data, project = "stickleback.sample.3")
cdna3[["percent.mt"]] <- PercentageFeatureSet(cdna3, pattern = "^mt-" )
cdna3 <- subset(cdna3, subset = nFeature_RNA > 100 & nFeature_RNA < 500 & percent.mt < 2)
VlnPlot(cdna3, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, pt.size = 0.1) 
plot1 <- FeatureScatter(cdna3, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(cdna3, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
CombinePlots(plots = list(plot1, plot2))
```

## Sample 4 QC

```{r fig.width=11}
cdna4 <- CreateSeuratObject(counts = cdna4.data, project = "stickleback.sample.4")
cdna4[["percent.mt"]] <- PercentageFeatureSet(cdna4, pattern = "^mt-"  )
cdna4 <- subset(cdna4, subset = nFeature_RNA > 100 & nFeature_RNA < 500 & percent.mt < 2)
VlnPlot(cdna4, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, pt.size = 0.1) 
plot1 <- FeatureScatter(cdna4, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(cdna4, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
CombinePlots(plots = list(plot1, plot2))
```

```{r}
stickleback.rna <- merge(x = cdna1, y = c(cdna2, cdna3, cdna4), add.cell.ids = c("cDNA1", "cDNA2", "cDNA3", "cDNA4"), project = "stickleback.rna" )

saveRDS(stickleback.rna, file = "stickleback.rna.seurat.obj.filtered.rds")
# stickleback.rna <- readRDS(file = "stickleback.rna.seurat.obj.filtered.rds")

# stickleback.rna <- subset(stickleback.rna, subset = nFeature_RNA > 50 & nFeature_RNA < 1000)
rm(cdna1.data, cdna2.data, cdna3.data, cdna4.data, cdna1, cdna2, cdna3, cdna4)
```


```{r}
#=========================================== SEURAT QC ==============================================
# stickleback.rna[["percent.mt"]] <- PercentageFeatureSet(stickleback.rna, pattern = "^mt-",  )
# 
# VlnPlot(stickleback.rna, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
```

```{r fig.width=15, fig.height=12}
#=========================================== SEURAT QC ==============================================
VlnPlot(stickleback.rna, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, pt.size = 0.1) 
  
plot1 <- FeatureScatter(stickleback.rna, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(stickleback.rna, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
CombinePlots(plots = list(plot1, plot2))
# stickleback.rna <- subset(stickleback.rna, subset = nFeature_RNA > 110 & nFeature_RNA < 500 & percent.mt < 2)
```


### Highest variable genes

```{r fig.width=12, fig.height=9, warning=FALSE}
#=========================================== SEURAT PIPELINE ==============================================
stickleback.rna <- NormalizeData(stickleback.rna, normalization.method = "LogNormalize", scale.factor = 10000)

stickleback.rna <- FindVariableFeatures(stickleback.rna, selection.method = "vst", nfeatures = 2000)
# Identify the 10 most highly variable genes
top10 <- head(VariableFeatures(stickleback.rna), 10)

# plot variable features with and without labels
plot1 <- VariableFeaturePlot(stickleback.rna)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
CombinePlots(plots = list(plot1, plot2))
```

```{r}
all.genes <- rownames(stickleback.rna)
stickleback.rna <- ScaleData(stickleback.rna, features = all.genes)
```

### PCA plot

```{r}
stickleback.rna <- RunPCA(stickleback.rna, features = VariableFeatures(object = stickleback.rna))
DimPlot(stickleback.rna, reduction = "pca")

```

### Approximate dimensionality of the dataset

```{r}
ElbowPlot(stickleback.rna)
```


```{r}
stickleback.rna <- FindNeighbors(stickleback.rna, dims = 1:15)
stickleback.rna <- RunUMAP(stickleback.rna, dims = 1:15)
saveRDS(stickleback.rna, file = "stickleback.rna.after.clustering.obj.rds")
```

### Clusterisation with low resolution (resolution = 0.06)
```{r}
stickleback.rna <- FindClusters(stickleback.rna, resolution = 0.06)
DimPlot(stickleback.rna, reduction = "umap")
```

### Clusterisation with high resolution (resolution = 0.3)
```{r}
stickleback.rna <- FindClusters(stickleback.rna, resolution = 0.3)
DimPlot(stickleback.rna, reduction = "umap")
```

### Distribution of cells by sample

```{r}
DimPlot(stickleback.rna, reduction = "umap", group.by = "orig.ident")
```




```{r}
#======================================== SMAPLE BATCH CORRECTION =============================================
# Should not be attempted before filtration step due to RAM consumption!!
stickleback.rna.comb <- readRDS(file = "stickleback.rna.seurat.obj.filtered.rds")
rna.list <- SplitObject(stickleback.rna.comb, split.by = "orig.ident")
for (i in 1:length(rna.list)) {
    rna.list[[i]] <- NormalizeData(rna.list[[i]], verbose = FALSE)
    rna.list[[i]] <- FindVariableFeatures(rna.list[[i]], selection.method = "vst", 
        nfeatures = 2000, verbose = FALSE)
    # rna.list[[i]] <- subset(rna.list[[i]], subset = nFeature_RNA > 200 & nFeature_RNA < 2500) #Filtration step, essential!!
}

rna.anchors <- FindIntegrationAnchors(object.list = rna.list, dims = 1:20)
rna.integrated <- IntegrateData(anchorset = rna.anchors, dims = 1:30)


```

## Integrated all 4 samples with batch correction

```{r}
#======================================== INTEGRATED DATA ANALYSIS =============================================
if(!exists("rna.integrated")){
  load(file = "rna.integrated.RData")
}
# switch to integrated assay. The variable features of this assay are automatically
# set during IntegrateData
DefaultAssay(rna.integrated) <- "integrated"

# Run the standard workflow for visualization and clustering
rna.integrated <- ScaleData(rna.integrated, verbose = FALSE)
rna.integrated <- RunPCA(rna.integrated, npcs = 20, verbose = FALSE)
rna.integrated <- RunUMAP(rna.integrated, reduction = "pca", dims = 1:20)
p1 <- DimPlot(rna.integrated, reduction = "umap")
# p2 <- DimPlot(rna.integrated, reduction = "umap", group.by = "celltype", label = TRUE, 
#     repel = TRUE) + NoLegend()
# plot_grid(p1, p2)
p1
```

### Clusterisation of integrated data with low resolution (resolution = 0.06)

```{r}
rna.integrated <- FindNeighbors(rna.integrated, dims = 1:15)
rna.integrated <- FindClusters(rna.integrated, resolution = 0.06)
rna.integrated <- RunUMAP(rna.integrated, dims = 1:15)
DimPlot(rna.integrated, reduction = "umap")

```

### Clusterisation of integrated data with higher resolution (resolution = 0.1)

```{r}
rna.integrated <- FindClusters(rna.integrated, resolution = 0.1)
DimPlot(rna.integrated, reduction = "umap")
save(rna.integrated, file = "rna.integrated.filtered.after.clustering.obj.RData")
```


### Top markers for each cluster, after batch correction (resolution = 0.1)
```{r}
#======================================== CUSTER MARKERS =============================================
# RESOLUTION = 0.1
rna.integrated <- FindClusters(rna.integrated, resolution = 0.1)
markers_0.1.rna.integrated  <- FindAllMarkers(rna.integrated, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.15)
write.csv(markers_0.1.rna.integrated, file = "stickleback.rna.batch.corrected.res_0.1_markers.csv")
# markers_0.1.rna.integrated <- read.csv(file = "stickleback.rna.batch.corrected.res_0.1_markers.csv")

#top for each cluster
markers_0.1.rna.integrated %>% group_by(cluster) %>% select(avg_logFC, p_val_adj, cluster, gene) %>% top_n(n = 4, wt = avg_logFC)
```

```{r}
#=================================== ID FIX ==========================================
# load(file = "rna.integrated.filtered.after.clustering.obj.RData")
mixed.ids <- data.frame(
  index = 1:length(rownames(rna.integrated)),
  mixed = rownames(rna.integrated))
hgnc.ids <- mixed.ids[grep("^ENSGACG", mixed.ids$mixed, invert = TRUE),]

raw_cdna1 <- Read10X(data.dir = "cdna1_counts/", gene.column = 1)
cdna1.features <- read.delim(file = "cdna1_counts/features.tsv.gz", col.names = c("ensembl_id", "mixed_id", "type"))

```

```{r}
#=================================== Count UMIs per cluster ==========================================
load(file = "rna.integrated.filtered.after.clustering.obj.RData")
rna.integrated.cluster.1 <- subset(rna.integrated, idents = 1)
rna.integrated.cluster.2 <- subset(rna.integrated, idents = 2)
```