---
title: "Zebrafish vs Stickleback pseudobulk comparisons"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(echo=FALSE, message = FALSE)
```

```{r, results = 'hide'}
library(ggplot2)
library(dplyr)
library(Seurat)
```

```{r separating_blood_cluster}
load(file = "data/rna.integrated.filtered.after.clustering.obj.RData")
rna.integrated.cluster.blood <- subset(rna.integrated, idents = 0)
rna.integrated.cluster.rest <- subset(rna.integrated, idents = c(1,2))
# save(rna.integrated.cluster.blood, file = "data/rna.integrated.cluster.blood.RData")
# save(rna.integrated.cluster.rest, file = "data/rna.integrated.cluster.rest.RData")
```


```{r importing_zebrafish_data}
# zebrafish_obj <- readRDS(file = "data/Farnsworth_2019.rds")
#     # zebrafish_gills <- subset(zebrafish_obj, idents = 5)
#     # Create dataframe with cluster numbers and cell names
# meta_table <- as.data.frame(zebrafish_obj@ident)
# meta_table$cell_names <- rownames(meta_table)
# meta_table$cluster <- as.character(meta_table$`zebrafish_obj@ident`)
# meta_table$`zebrafish_obj@ident` <- NULL
# meta_gills <- subset(meta_table, cluster %in% c(5, 21, 146))
#     # extract counts table from the seurat obj
# zebrafish_counts <- zebrafish_obj@raw.data
#     # subset counts as regular matrix
# gill_counts <- as.data.frame(zebrafish_counts[, colnames(zebrafish_counts) %in% meta_gills$cell_names])
```



### Zebrafish vs non-blood Stickleback clusters (5, 21, 146 marked as Gill clusters a, b, c)


```{r converting_data_to_bulk_rest_stickleback}
# converting zebrafish
# gill_bulk_zebrafish <- as.data.frame(apply(gill_counts, 1, sum))
# colnames(gill_bulk_zebrafish) <- c("summed_counts_zebrafish")
# gill_bulk_zebrafish$gene <- rownames(gill_bulk_zebrafish)
# write.table(gill_bulk_zebrafish, file = "data/gill_bulk_zebrafish.tsv", sep = "\t")
gill_bulk_zebrafish <- read.table(file = "data/gill_bulk_zebrafish.tsv")

# converting stickleback ONLY REST
gill_counts_stickleback_rest <- as.data.frame(rna.integrated.cluster.rest@assays$RNA@counts)
gill_bulk_stickleback_rest <- as.data.frame(apply(gill_counts_stickleback_rest, 1, sum))
gill_bulk_stickleback_rest$gene <- rownames(gill_bulk_stickleback_rest)

# subsetting only common ids, probably vastly not optimal since I have not seen ENSEMBL ids in the zebrafish dataset, unlike the stickleback
# ONLY HGNC WERE LEFT AFTER THE MERGER !!!!
common_ids <- intersect(rownames(gill_bulk_stickleback_rest), rownames(gill_bulk_zebrafish))
gill_bulk_stickleback_rest <- subset(gill_bulk_stickleback_rest, gene %in% common_ids)
gill_bulk_zebrafish <- subset(gill_bulk_zebrafish, gene %in% common_ids)

# combining in a single table for the scatterplot
gill_bulk_both <- merge(gill_bulk_stickleback_rest, gill_bulk_zebrafish, by = "gene")
colnames(gill_bulk_both) <- c("gene", "stickleback_counts", "zebrafish_counts")
```

```{r scatter_plot_comparing_stickleback_and_zebrafish_pseudobulk_REST}
require(ggpubr)

gill_bulk_both_log <- transform(gill_bulk_both, stickleback_counts = log(stickleback_counts), zebrafish_counts = log(zebrafish_counts))

ggscatter(gill_bulk_both_log,
      x = "stickleback_counts", y = "zebrafish_counts", size = 1.5) +
      theme_gray() +
      stat_cor(method = "pearson", label.x = -5, label.y = 5, size=3, color="#696969")+
      # coord_cartesian(xlim = c(-5, 5), ylim = c(-5, 5)) +
      labs(title = "Corelation scatter plot pseudobulk stickleback vs zebrafish")
```

Non-blood clusters show bigger correlation with gill cells by expression, as expected.


### Zebrafish vs blood Stickleback clusters (5, 21, 146 marked as Gill clusters a, b, c)

```{r converting_data_to_bulk_blood_stickleback}
# converting zebrafish
gill_bulk_zebrafish <- as.data.frame(apply(gill_counts, 1, sum))
colnames(gill_bulk_zebrafish) <- c("summed_counts_zebrafish")
gill_bulk_zebrafish$gene <- rownames(gill_bulk_zebrafish)

# converting stickleback ONLY BLOOD
gill_counts_stickleback_blood <- as.data.frame(rna.integrated.cluster.blood@assays$RNA@counts)
gill_bulk_stickleback_blood <- as.data.frame(apply(gill_counts_stickleback_blood, 1, sum))
gill_bulk_stickleback_blood$gene <- rownames(gill_bulk_stickleback_blood)

# subsetting only common ids, probably vastly not optimal since I have not seen ENSEMBL ids in the zebrafish dataset, unlike the stickleback
# ONLY HGNC WERE LEFT AFTER THE MERGER !!!!
common_ids <- intersect(rownames(gill_bulk_stickleback_blood), rownames(gill_bulk_zebrafish))
gill_bulk_stickleback_blood <- subset(gill_bulk_stickleback_blood, gene %in% common_ids)
gill_bulk_zebrafish <- subset(gill_bulk_zebrafish, gene %in% common_ids)

# combining in a single table for the scatterplot
gill_bulk_both <- merge(gill_bulk_stickleback_blood, gill_bulk_zebrafish, by = "gene")
colnames(gill_bulk_both) <- c("gene", "stickleback_counts", "zebrafish_counts")
```

```{r scatter_plot_comparing_stickleback_and_zebrafish_pseudobulk_BLOOD}
require(ggpubr)

gill_bulk_both_log <- transform(gill_bulk_both, stickleback_counts = log(stickleback_counts), zebrafish_counts = log(zebrafish_counts))

ggscatter(gill_bulk_both_log,
      x = "stickleback_counts", y = "zebrafish_counts", size = 1.5) +
      theme_gray() +
      stat_cor(method = "pearson", label.x = -5, label.y = 5, size=3, color="#696969")+
      # coord_cartesian(xlim = c(-5, 5), ylim = c(-5, 5)) +
      labs(title = "Corelation scatter plot pseudobulk stickleback vs zebrafish")
```


```{r zebrafish_clusters_separately}
zebrafish_obj <- readRDS(file = "data/Farnsworth_2019.rds")
    # zebrafish_gills <- subset(zebrafish_obj, idents = 5)
    # Create dataframe with cluster numbers and cell names
meta_table <- as.data.frame(zebrafish_obj@ident)
meta_table$cell_names <- rownames(meta_table)
meta_table$cluster <- as.character(meta_table$`zebrafish_obj@ident`)
meta_table$`zebrafish_obj@ident` <- NULL
meta_gills <- subset(meta_table, cluster %in% c(5, 21, 146, 55, 71, 60, 7))

meta_gills_5 <- subset(meta_gills, cluster == 5)
meta_gills_21 <- subset(meta_gills, cluster == 21)
meta_gills_146 <- subset(meta_gills, cluster == 146)
    # extract counts table from the seurat obj
zebrafish_counts <- zebrafish_obj@raw.data
    # subset counts as regular matrix
gill_counts_1 <- as.data.frame(zebrafish_counts[, colnames(zebrafish_counts) %in% meta_gills_55$cell_names])
gill_counts_5 <- as.data.frame(zebrafish_counts[, colnames(zebrafish_counts) %in% meta_gills_5$cell_names])
gill_counts_21 <- as.data.frame(zebrafish_counts[, colnames(zebrafish_counts) %in% meta_gills_21$cell_names])
gill_counts_146 <- as.data.frame(zebrafish_counts[, colnames(zebrafish_counts) %in% meta_gills_146$cell_names])
   # converting zebrafish
gill_bulk_zebrafish_cluster_5 <- as.data.frame(apply(gill_counts_5, 1, sum))
colnames(gill_bulk_zebrafish_cluster_5) <- c("summed_counts_zebrafish")
gill_bulk_zebrafish_cluster_5$gene <- rownames(gill_bulk_zebrafish_cluster_5)
write.table(gill_bulk_zebrafish_cluster_5, file = "data/gill_bulk_zebrafish_cluster_5.tsv", sep = "\t")

gill_bulk_zebrafish_cluster_21 <- as.data.frame(apply(gill_counts_21, 1, sum))
colnames(gill_bulk_zebrafish_cluster_21) <- c("summed_counts_zebrafish")
gill_bulk_zebrafish_cluster_21$gene <- rownames(gill_bulk_zebrafish_cluster_21)
write.table(gill_bulk_zebrafish_cluster_21, file = "data/gill_bulk_zebrafish_cluster_21.tsv", sep = "\t")

gill_bulk_zebrafish_cluster_146 <- as.data.frame(apply(gill_counts_146, 1, sum))
colnames(gill_bulk_zebrafish_cluster_146) <- c("summed_counts_zebrafish")
gill_bulk_zebrafish_cluster_146$gene <- rownames(gill_bulk_zebrafish_cluster_146)
write.table(gill_bulk_zebrafish_cluster_146, file = "data/gill_bulk_zebrafish_cluster_146.tsv", sep = "\t")





gill_bulk_zebrafish_cluster_5 <- read.table(file = "data/gill_bulk_zebrafish_cluster_5.tsv")
gill_bulk_zebrafish_cluster_21 <- read.table(file = "data/gill_bulk_zebrafish_cluster_21.tsv")
gill_bulk_zebrafish_cluster_146 <- read.table(file = "data/gill_bulk_zebrafish_cluster_146.tsv")

meta_gills_55 <- subset(meta_gills, cluster == 55)
gill_counts_55 <- as.data.frame(zebrafish_counts[, colnames(zebrafish_counts) %in% meta_gills_55$cell_names])
gill_bulk_zebrafish_cluster_55 <- as.data.frame(apply(gill_counts_55, 1, sum))
colnames(gill_bulk_zebrafish_cluster_55) <- c("summed_counts_zebrafish")
gill_bulk_zebrafish_cluster_55$gene <- rownames(gill_bulk_zebrafish_cluster_55)

meta_gills_71 <- subset(meta_gills, cluster == 71)
gill_counts_71 <- as.data.frame(zebrafish_counts[, colnames(zebrafish_counts) %in% meta_gills_71$cell_names])
gill_bulk_zebrafish_cluster_71 <- as.data.frame(apply(gill_counts_71, 1, sum))
colnames(gill_bulk_zebrafish_cluster_71) <- c("summed_counts_zebrafish")
gill_bulk_zebrafish_cluster_71$gene <- rownames(gill_bulk_zebrafish_cluster_71)

meta_gills_60 <- subset(meta_gills, cluster == 60)
gill_counts_60 <- as.data.frame(zebrafish_counts[, colnames(zebrafish_counts) %in% meta_gills_60$cell_names])
gill_bulk_zebrafish_cluster_60 <- as.data.frame(apply(gill_counts_60, 1, sum))
colnames(gill_bulk_zebrafish_cluster_60) <- c("summed_counts_zebrafish")
gill_bulk_zebrafish_cluster_60$gene <- rownames(gill_bulk_zebrafish_cluster_60)

meta_gills_7 <- subset(meta_gills, cluster == 7)
gill_counts_7 <- as.data.frame(zebrafish_counts[, colnames(zebrafish_counts) %in% meta_gills_7$cell_names])
gill_bulk_zebrafish_cluster_7 <- as.data.frame(apply(gill_counts_7, 1, sum))
colnames(gill_bulk_zebrafish_cluster_7) <- c("summed_counts_zebrafish")
gill_bulk_zebrafish_cluster_7$gene <- rownames(gill_bulk_zebrafish_cluster_7)
```

### 1 Zebrafish gill cluster vs non-blood Stickleback clusters (5 marked as Gill clusters a)


```{r converting_data_to_bulk_rest_stickleback_zebrafish_cluster_5}
# converting zebrafish
# gill_bulk_zebrafish <- as.data.frame(apply(gill_counts, 1, sum))
# colnames(gill_bulk_zebrafish) <- c("summed_counts_zebrafish")
# gill_bulk_zebrafish$gene <- rownames(gill_bulk_zebrafish)
# write.table(gill_bulk_zebrafish, file = "data/gill_bulk_zebrafish.tsv", sep = "\t")
gill_bulk_zebrafish <- gill_bulk_zebrafish_cluster_5

# converting stickleback ONLY REST
gill_counts_stickleback_rest <- as.data.frame(rna.integrated.cluster.rest@assays$RNA@counts)
gill_bulk_stickleback_rest <- as.data.frame(apply(gill_counts_stickleback_rest, 1, sum))
gill_bulk_stickleback_rest$gene <- rownames(gill_bulk_stickleback_rest)

# subsetting only common ids, probably vastly not optimal since I have not seen ENSEMBL ids in the zebrafish dataset, unlike the stickleback
# ONLY HGNC WERE LEFT AFTER THE MERGER !!!!
common_ids <- intersect(rownames(gill_bulk_stickleback_rest), rownames(gill_bulk_zebrafish))
gill_bulk_stickleback_rest <- subset(gill_bulk_stickleback_rest, gene %in% common_ids)
gill_bulk_zebrafish <- subset(gill_bulk_zebrafish, gene %in% common_ids)

# combining in a single table for the scatterplot
gill_bulk_both <- merge(gill_bulk_stickleback_rest, gill_bulk_zebrafish, by = "gene")
colnames(gill_bulk_both) <- c("gene", "stickleback_counts", "zebrafish_counts")
```

```{r scatter_plot_comparing_stickleback_and_zebrafish_pseudobulk_REST_zebrafish_cluster_5}
require(ggpubr)

gill_bulk_both_log <- transform(gill_bulk_both, stickleback_counts = log(stickleback_counts), zebrafish_counts = log(zebrafish_counts))

ggscatter(gill_bulk_both_log,
      x = "stickleback_counts", y = "zebrafish_counts", size = 1.5) +
      theme_gray() +
      stat_cor(method = "pearson", label.x = -5, label.y = 5, size=3, color="#696969")+
      # coord_cartesian(xlim = c(-5, 5), ylim = c(-5, 5)) +
      labs(title = "Corelation scatter plot pseudobulk stickleback vs zebrafish cluster 5")
```

### 1 Zebrafish gill cluster vs non-blood Stickleback clusters (21 marked as Gill clusters b)


```{r converting_data_to_bulk_rest_stickleback_zebrafish_cluster_21}
# converting zebrafish
# gill_bulk_zebrafish <- as.data.frame(apply(gill_counts, 1, sum))
# colnames(gill_bulk_zebrafish) <- c("summed_counts_zebrafish")
# gill_bulk_zebrafish$gene <- rownames(gill_bulk_zebrafish)
# write.table(gill_bulk_zebrafish, file = "data/gill_bulk_zebrafish.tsv", sep = "\t")
gill_bulk_zebrafish <- gill_bulk_zebrafish_cluster_21

# converting stickleback ONLY REST
gill_counts_stickleback_rest <- as.data.frame(rna.integrated.cluster.rest@assays$RNA@counts)
gill_bulk_stickleback_rest <- as.data.frame(apply(gill_counts_stickleback_rest, 1, sum))
gill_bulk_stickleback_rest$gene <- rownames(gill_bulk_stickleback_rest)

# subsetting only common ids, probably vastly not optimal since I have not seen ENSEMBL ids in the zebrafish dataset, unlike the stickleback
# ONLY HGNC WERE LEFT AFTER THE MERGER !!!!
common_ids <- intersect(rownames(gill_bulk_stickleback_rest), rownames(gill_bulk_zebrafish))
gill_bulk_stickleback_rest <- subset(gill_bulk_stickleback_rest, gene %in% common_ids)
gill_bulk_zebrafish <- subset(gill_bulk_zebrafish, gene %in% common_ids)

# combining in a single table for the scatterplot
gill_bulk_both <- merge(gill_bulk_stickleback_rest, gill_bulk_zebrafish, by = "gene")
colnames(gill_bulk_both) <- c("gene", "stickleback_counts", "zebrafish_counts")
```

```{r scatter_plot_comparing_stickleback_and_zebrafish_pseudobulk_REST_zebrafish_cluster_21}
require(ggpubr)

gill_bulk_both_log <- transform(gill_bulk_both, stickleback_counts = log(stickleback_counts), zebrafish_counts = log(zebrafish_counts))

ggscatter(gill_bulk_both_log,
      x = "stickleback_counts", y = "zebrafish_counts", size = 1.5) +
      theme_gray() +
      stat_cor(method = "pearson", label.x = -5, label.y = 5, size=3, color="#696969")+
      # coord_cartesian(xlim = c(-5, 5), ylim = c(-5, 5)) +
      labs(title = "Corelation scatter plot pseudobulk stickleback vs zebrafish cluster 21")
```

### 1 Zebrafish gill cluster vs non-blood Stickleback clusters (146 marked as Gill clusters c)


```{r converting_data_to_bulk_rest_stickleback_zebrafish_cluster_146}
# converting zebrafish
# gill_bulk_zebrafish <- as.data.frame(apply(gill_counts, 1, sum))
# colnames(gill_bulk_zebrafish) <- c("summed_counts_zebrafish")
# gill_bulk_zebrafish$gene <- rownames(gill_bulk_zebrafish)
# write.table(gill_bulk_zebrafish, file = "data/gill_bulk_zebrafish.tsv", sep = "\t")
gill_bulk_zebrafish <- gill_bulk_zebrafish_cluster_146

# converting stickleback ONLY REST
gill_counts_stickleback_rest <- as.data.frame(rna.integrated.cluster.rest@assays$RNA@counts)
gill_bulk_stickleback_rest <- as.data.frame(apply(gill_counts_stickleback_rest, 1, sum))
gill_bulk_stickleback_rest$gene <- rownames(gill_bulk_stickleback_rest)

# subsetting only common ids, probably vastly not optimal since I have not seen ENSEMBL ids in the zebrafish dataset, unlike the stickleback
# ONLY HGNC WERE LEFT AFTER THE MERGER !!!!
common_ids <- intersect(rownames(gill_bulk_stickleback_rest), rownames(gill_bulk_zebrafish))
gill_bulk_stickleback_rest <- subset(gill_bulk_stickleback_rest, gene %in% common_ids)
gill_bulk_zebrafish <- subset(gill_bulk_zebrafish, gene %in% common_ids)

# combining in a single table for the scatterplot
gill_bulk_both <- merge(gill_bulk_stickleback_rest, gill_bulk_zebrafish, by = "gene")
colnames(gill_bulk_both) <- c("gene", "stickleback_counts", "zebrafish_counts")
```

```{r scatter_plot_comparing_stickleback_and_zebrafish_pseudobulk_REST_zebrafish_cluster_146}
require(ggpubr)

gill_bulk_both_log <- transform(gill_bulk_both, stickleback_counts = log(stickleback_counts), zebrafish_counts = log(zebrafish_counts))

ggscatter(gill_bulk_both_log,
      x = "stickleback_counts", y = "zebrafish_counts", size = 1.5) +
      theme_gray() +
      stat_cor(method = "pearson", label.x = -5, label.y = 5, size=3, color="#696969")+
      # coord_cartesian(xlim = c(-5, 5), ylim = c(-5, 5)) +
      labs(title = "Corelation scatter plot pseudobulk stickleback vs zebrafish cluster 146")
```



### 1 Zebrafish gill cluster vs non-blood Stickleback clusters (55 marked as Liver cells)


```{r converting_data_to_bulk_rest_stickleback_zebrafish_cluster_55}
# converting zebrafish
# gill_bulk_zebrafish <- as.data.frame(apply(gill_counts, 1, sum))
# colnames(gill_bulk_zebrafish) <- c("summed_counts_zebrafish")
# gill_bulk_zebrafish$gene <- rownames(gill_bulk_zebrafish)
# write.table(gill_bulk_zebrafish, file = "data/gill_bulk_zebrafish.tsv", sep = "\t")
gill_bulk_zebrafish <- gill_bulk_zebrafish_cluster_55

# converting stickleback ONLY REST
gill_counts_stickleback_rest <- as.data.frame(rna.integrated.cluster.rest@assays$RNA@counts)
gill_bulk_stickleback_rest <- as.data.frame(apply(gill_counts_stickleback_rest, 1, sum))
gill_bulk_stickleback_rest$gene <- rownames(gill_bulk_stickleback_rest)

# subsetting only common ids, probably vastly not optimal since I have not seen ENSEMBL ids in the zebrafish dataset, unlike the stickleback
# ONLY HGNC WERE LEFT AFTER THE MERGER !!!!
common_ids <- intersect(rownames(gill_bulk_stickleback_rest), rownames(gill_bulk_zebrafish))
gill_bulk_stickleback_rest <- subset(gill_bulk_stickleback_rest, gene %in% common_ids)
gill_bulk_zebrafish <- subset(gill_bulk_zebrafish, gene %in% common_ids)

# combining in a single table for the scatterplot
gill_bulk_both <- merge(gill_bulk_stickleback_rest, gill_bulk_zebrafish, by = "gene")
colnames(gill_bulk_both) <- c("gene", "stickleback_counts", "zebrafish_counts")
```

```{r scatter_plot_comparing_stickleback_and_zebrafish_pseudobulk_REST_zebrafish_cluster_55}
require(ggpubr)

gill_bulk_both_log <- transform(gill_bulk_both, stickleback_counts = log(stickleback_counts), zebrafish_counts = log(zebrafish_counts))

ggscatter(gill_bulk_both_log,
      x = "stickleback_counts", y = "zebrafish_counts", size = 1.5) +
      theme_gray() +
      stat_cor(method = "pearson", label.x = -5, label.y = 5, size=3, color="#696969")+
      # coord_cartesian(xlim = c(-5, 5), ylim = c(-5, 5)) +
      labs(title = "Corelation scatter plot pseudobulk stickleback vs zebrafish cluster 55")
```

### 1 Zebrafish gill cluster vs non-blood Stickleback clusters (71 marked as Macrophage cells)


```{r converting_data_to_bulk_rest_stickleback_zebrafish_cluster_71}
# converting zebrafish
# gill_bulk_zebrafish <- as.data.frame(apply(gill_counts, 1, sum))
# colnames(gill_bulk_zebrafish) <- c("summed_counts_zebrafish")
# gill_bulk_zebrafish$gene <- rownames(gill_bulk_zebrafish)
# write.table(gill_bulk_zebrafish, file = "data/gill_bulk_zebrafish.tsv", sep = "\t")
gill_bulk_zebrafish <- gill_bulk_zebrafish_cluster_71

# converting stickleback ONLY REST
gill_counts_stickleback_rest <- as.data.frame(rna.integrated.cluster.rest@assays$RNA@counts)
gill_bulk_stickleback_rest <- as.data.frame(apply(gill_counts_stickleback_rest, 1, sum))
gill_bulk_stickleback_rest$gene <- rownames(gill_bulk_stickleback_rest)

# subsetting only common ids, probably vastly not optimal since I have not seen ENSEMBL ids in the zebrafish dataset, unlike the stickleback
# ONLY HGNC WERE LEFT AFTER THE MERGER !!!!
common_ids <- intersect(rownames(gill_bulk_stickleback_rest), rownames(gill_bulk_zebrafish))
gill_bulk_stickleback_rest <- subset(gill_bulk_stickleback_rest, gene %in% common_ids)
gill_bulk_zebrafish <- subset(gill_bulk_zebrafish, gene %in% common_ids)

# combining in a single table for the scatterplot
gill_bulk_both <- merge(gill_bulk_stickleback_rest, gill_bulk_zebrafish, by = "gene")
colnames(gill_bulk_both) <- c("gene", "stickleback_counts", "zebrafish_counts")
```

```{r scatter_plot_comparing_stickleback_and_zebrafish_pseudobulk_REST_zebrafish_cluster_71}
require(ggpubr)

gill_bulk_both_log <- transform(gill_bulk_both, stickleback_counts = log(stickleback_counts), zebrafish_counts = log(zebrafish_counts))

ggscatter(gill_bulk_both_log,
      x = "stickleback_counts", y = "zebrafish_counts", size = 1.5) +
      theme_gray() +
      stat_cor(method = "pearson", label.x = -5, label.y = 5, size=3, color="#696969")+
      # coord_cartesian(xlim = c(-5, 5), ylim = c(-5, 5)) +
      labs(title = "Corelation scatter plot pseudobulk stickleback vs zebrafish cluster 71")
```

### 1 Zebrafish gill cluster vs non-blood Stickleback clusters (60 marked as Retina cells)


```{r converting_data_to_bulk_rest_stickleback_zebrafish_cluster_60}
# converting zebrafish
# gill_bulk_zebrafish <- as.data.frame(apply(gill_counts, 1, sum))
# colnames(gill_bulk_zebrafish) <- c("summed_counts_zebrafish")
# gill_bulk_zebrafish$gene <- rownames(gill_bulk_zebrafish)
# write.table(gill_bulk_zebrafish, file = "data/gill_bulk_zebrafish.tsv", sep = "\t")
gill_bulk_zebrafish <- gill_bulk_zebrafish_cluster_60

# converting stickleback ONLY REST
gill_counts_stickleback_rest <- as.data.frame(rna.integrated.cluster.rest@assays$RNA@counts)
gill_bulk_stickleback_rest <- as.data.frame(apply(gill_counts_stickleback_rest, 1, sum))
gill_bulk_stickleback_rest$gene <- rownames(gill_bulk_stickleback_rest)

# subsetting only common ids, probably vastly not optimal since I have not seen ENSEMBL ids in the zebrafish dataset, unlike the stickleback
# ONLY HGNC WERE LEFT AFTER THE MERGER !!!!
common_ids <- intersect(rownames(gill_bulk_stickleback_rest), rownames(gill_bulk_zebrafish))
gill_bulk_stickleback_rest <- subset(gill_bulk_stickleback_rest, gene %in% common_ids)
gill_bulk_zebrafish <- subset(gill_bulk_zebrafish, gene %in% common_ids)

# combining in a single table for the scatterplot
gill_bulk_both <- merge(gill_bulk_stickleback_rest, gill_bulk_zebrafish, by = "gene")
colnames(gill_bulk_both) <- c("gene", "stickleback_counts", "zebrafish_counts")
```

```{r scatter_plot_comparing_stickleback_and_zebrafish_pseudobulk_REST_zebrafish_cluster_60}
require(ggpubr)

gill_bulk_both_log <- transform(gill_bulk_both, stickleback_counts = log(stickleback_counts), zebrafish_counts = log(zebrafish_counts))

ggscatter(gill_bulk_both_log,
      x = "stickleback_counts", y = "zebrafish_counts", size = 1.5) +
      theme_gray() +
      stat_cor(method = "pearson", label.x = -5, label.y = 5, size=3, color="#696969")+
      # coord_cartesian(xlim = c(-5, 5), ylim = c(-5, 5)) +
      labs(title = "Corelation scatter plot pseudobulk stickleback vs zebrafish cluster 60")
```

### 1 Zebrafish gill cluster vs non-blood Stickleback clusters (7 marked as Muscle cells)


```{r converting_data_to_bulk_rest_stickleback_zebrafish_cluster_7}
# converting zebrafish
# gill_bulk_zebrafish <- as.data.frame(apply(gill_counts, 1, sum))
# colnames(gill_bulk_zebrafish) <- c("summed_counts_zebrafish")
# gill_bulk_zebrafish$gene <- rownames(gill_bulk_zebrafish)
# write.table(gill_bulk_zebrafish, file = "data/gill_bulk_zebrafish.tsv", sep = "\t")
gill_bulk_zebrafish <- gill_bulk_zebrafish_cluster_7

# converting stickleback ONLY REST
gill_counts_stickleback_rest <- as.data.frame(rna.integrated.cluster.rest@assays$RNA@counts)
gill_bulk_stickleback_rest <- as.data.frame(apply(gill_counts_stickleback_rest, 1, sum))
gill_bulk_stickleback_rest$gene <- rownames(gill_bulk_stickleback_rest)

# subsetting only common ids, probably vastly not optimal since I have not seen ENSEMBL ids in the zebrafish dataset, unlike the stickleback
# ONLY HGNC WERE LEFT AFTER THE MERGER !!!!
common_ids <- intersect(rownames(gill_bulk_stickleback_rest), rownames(gill_bulk_zebrafish))
gill_bulk_stickleback_rest <- subset(gill_bulk_stickleback_rest, gene %in% common_ids)
gill_bulk_zebrafish <- subset(gill_bulk_zebrafish, gene %in% common_ids)

# combining in a single table for the scatterplot
gill_bulk_both <- merge(gill_bulk_stickleback_rest, gill_bulk_zebrafish, by = "gene")
colnames(gill_bulk_both) <- c("gene", "stickleback_counts", "zebrafish_counts")
```

```{r scatter_plot_comparing_stickleback_and_zebrafish_pseudobulk_REST_zebrafish_cluster_7}
require(ggpubr)

gill_bulk_both_log <- transform(gill_bulk_both, stickleback_counts = log(stickleback_counts), zebrafish_counts = log(zebrafish_counts))

ggscatter(gill_bulk_both_log,
      x = "stickleback_counts", y = "zebrafish_counts", size = 1.5) +
      theme_gray() +
      stat_cor(method = "pearson", label.x = -5, label.y = 5, size=3, color="#696969")+
      # coord_cartesian(xlim = c(-5, 5), ylim = c(-5, 5)) +
      labs(title = "Corelation scatter plot pseudobulk stickleback vs zebrafish cluster 7")
```


